### 2017-06-04 monocle_analysis.Rmd
Plot total mRNAs for infected and un-infected cells

---
title: "Monocle analysis of influenza-infected cells"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

## Setup
Load or install necessary packages.
If `Monocle` is not already available, install it using `Bioconductor`.
Note that Cole suggested that we might want to install the development version of GitHub; right now this does **not** do this.
```{r, message=FALSE, warning=FALSE}

if (!require("monocle")) {
  source("http://bioconductor.org/biocLite.R")
  biocLite()
  biocLite("monocle")
  }

if (!require("ggExtra")) install.packages("ggExtra")
if (!require("cowplot")) install.packages("cowplot")
```
```{r}
cat("Using Monocle version", as.character(packageVersion("monocle")), "\n")
```

## Create cell-gene dataset and examine some basic statistics about the number of influenza and cellular reads per cell
Read in cell-gene data of all samples.
This is the merged cell-gene data set constructed essentially from `cellranger` using the iPython notebook.
The cells are annotated by the number of wildtype and synonymous-barcoded influenza UMIs called for each cell.
```{r}
matrixfile <- "results/cellgenecounts/merged_humanplusflu_matrix.mtx"
genesfile <- "results/cellgenecounts/merged_humanplusflu_genes.tsv"
cellsfile <- "results/cellgenecounts/merged_humanplusflu_cells.tsv"

all_cells <- newCellDataSet(readMM(matrixfile), 
      phenoData=new("AnnotatedDataFrame", data=read.delim(cellsfile)), 
      featureData=new("AnnotatedDataFrame", data=read.delim(genesfile)),
      expressionFamily=negbinomial.size())
```

Filter on just the samples that interest us for this analysis.
Right now this is all the samples in the cell-gene dataset.
```{r}
samples = c("Uninfected", "6hr-1", "8hr-2", "8hr-2repeat", "10hr-2")

cellfilter <- row.names(subset(pData(all_cells), Sample %in% samples))
cells <- all_cells[,cellfilter]

cat("Filtered all", nrow(pData(all_cells)), "cells down to the", nrow(pData(cells)), 
      "in samples of interest.")
```

Compute the total number of mRNAs, cellular mRNAs, and influenza mRNAs per cell.
Also compute the fraction of total mRNAs that are from influenza.
```{r}
pData(cells)$total_mRNAs <- Matrix::colSums(exprs(cells))

flugenes <- c("fluPB2", "fluPB1", "fluPA", "fluHA", "fluNP", "fluNA", "fluM", "fluNS")
flugenes_id <- row.names(subset(fData(cells), gene_short_name %in% flugenes))
pData(cells)$flu_mRNAs <- Matrix::colSums(exprs(cells[flugenes_id,]))

pData(cells)$frac_flu_mRNAs <- pData(cells)$flu_mRNAs / pData(cells)$total_mRNAs

cellgenes_id <- row.names(subset(fData(cells), !(gene_short_name %in% flugenes)))
pData(cells)$cellular_mRNAs <- Matrix::colSums(exprs(cells[cellgenes_id,]))
```

Plot the number of influenza mRNAs versus the number of cellular RNAs as scatterplots.
The blue curves show the overall density on the number of cellular mRNAs.
Note that these numbers are not equivalent across samples as the different samples were sequenced to different depths.
```{r}
scatterplot <- ggplot(pData(cells), aes(cellular_mRNAs, flu_mRNAs)) + geom_point()

# next line uses eval substitute to get value of maxy as here: https://stackoverflow.com/a/10659563
densityplot <- eval(substitute(
    {stat_density(aes(x=cellular_mRNAs, y=maxy*(..scaled..)), geom="line", color='blue')}, 
    list(maxy=layer_scales(scatterplot)$y$range$range[2])))

scatterplot + densityplot + theme(axis.text.x=element_text(angle=90, vjust=0.5)) + facet_wrap(~Sample)
```

The `Monocle` vignette suggests possibly filtering out cells with far more / less mRNAs than most cells.
I have **not** done that for the reason that this will filter out quite a few of the influenza-infected cells.
This fact is made more clear by this next plot, which shows that the cells that have greater than 1% of their reads derived from influenza tend to have more total mRNAs than other cells in the two samples with the greatest number of influenza reads
I can think of several possible reasons for this trend: 
  1. Cells infected with influenza are simply "bigger" or are stimulated to produce more mRNAs
  2. Influenza infection preferentially leads to doublets due to cell clumping, perhaps mediated by hemagglutinin-based physical linkage of cells.
In any case, it seems safer **not** to exclude the influenza infected cells as we'd then be preferentially excluded cells with high viral burden.
```{r}
pData(cells)$frac_flu_gt_0.01 <- pData(cells)$frac_flu_mRNAs > 0.01

ggplot(pData(cells[, row.names(subset(pData(cells), Sample %in% c("8hr-2", "10hr-2")))]), 
        aes(total_mRNAs)) +
    geom_density(aes(color=frac_flu_gt_0.01)) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5)) +
    facet_wrap(~Sample)
#flugt100cells_id <- row.names(subset(pData(cells), flu_mRNAs > 50))
#flugt100_densityplot <- ggplot(pData(cells[,flugt100cells_id]), aes(total_mRNAs)) + geom_density()
```

```{r}
min_expr <- 5
num_cells <- 5
cells <- detectGenes(cells, min_expr=min_expr)
expressed_genes <- row.names(subset(fData(cells), num_cells_expressed >= num_cells))
cat("Filtering to the", length(expressed_genes), "genes expressed at >=", min_expr,
      "transcripts in at least", num_cells, "cells.\n")
cells <- cells[expressed_genes,]

cat("Plotting number of RNAs per cell for each sample\n")
print(table(pData(cells)$Sample))
pData(cells)$Total_mRNAs <- Matrix::colSums(exprs(cells))
qplot(Total_mRNAs, data=pData(cells), geom="density", color=Sample)

cat("Filtering out 6hr sample since it has a different distribution of RNA per cell")
cells <- cells[, row.names(subset(pData(cells), Sample != "6hr"))]
print(table(pData(cells)$Sample))
```

## Look at distribution of influenza genes.
```{r}
flu_id <- row.names(subset(fData(cells), substr(ShortGeneName, 1, 3) == "flu"))

pData(cells)$Total_flu <- Matrix::colSums(exprs(cells[flu_id,]))
qplot(Total_flu, data=pData(cells), color=Sample, log="x")

cth <- newCellTypeHierarchy()
cth <- addCellType(cth, "noFlu", classify_func=function(x) {colSums(x[flu_id,]) == 0})
cth <- addCellType(cth, "highFlu", classify_func=function(x) {colSums(x[flu_id,]) >= 100}) 
cells <- classifyCells(cells, cth)
print(table(pData(cells)$Sample, pData(cells)$CellType))
```