### 2017-06-05 monocle_analysis.Rmd
clean plots of basic flu vs cellular mRNAs

---
title: "Monocle analysis of influenza-infected cells"
author: "Alistair Russel, Cole Trapnell, Jesse Bloom"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

## Setup
Load or install necessary packages.
Note that Cole suggested that we might want to install the development version of [Monocle](http://cole-trapnell-lab.github.io/monocle-release/) from GitHub; right now this does **not** do this.
```{r, message=FALSE, warning=FALSE}

if (!require("monocle")) {
  source("http://bioconductor.org/biocLite.R")
  biocLite()
  biocLite("monocle")
  }

if (!require("ggExtra")) install.packages("ggExtra")
if (!require("cowplot")) install.packages("cowplot")
if (!require("scales")) install.packages("scales")

# scientific notation formatting: https://stackoverflow.com/a/24241954
fancy_scientific <- function(l) {
     l <- format(l, scientific=TRUE)
     l <- gsub("^0e\\+00","0", l)
     l <- gsub("^1e\\+00", "1", l)
     l <- gsub("^(.*)e", "'\\1'e", l)
     l <- gsub("e\\+","e", l)
     l <- gsub("e", "%*%10^", l)
     l <- gsub("^\'1\'\\%\\*\\%", "", l)
     parse(text=l)
}
```
```{r}
cat("Using Monocle version", as.character(packageVersion("monocle")), "\n")
```

## Create cell-gene dataset and examine basic statistics about number of influenza and cellular mRNAs per cell
Read in cell-gene data for all samples.
This is the merged cell-gene data set constructed with `cellranger` using the iPython notebook.
The cells are annotated by the number of wildtype and synonymous-barcoded influenza UMIs called for each cell.
```{r}
matrixfile <- "results/cellgenecounts/merged_humanplusflu_matrix.mtx"
genesfile <- "results/cellgenecounts/merged_humanplusflu_genes.tsv"
cellsfile <- "results/cellgenecounts/merged_humanplusflu_cells.tsv"

all_cells <- newCellDataSet(readMM(matrixfile), 
      phenoData=new("AnnotatedDataFrame", data=read.delim(cellsfile)), 
      featureData=new("AnnotatedDataFrame", data=read.delim(genesfile)),
      expressionFamily=negbinomial.size())
```

Filter on the samples that interest us.
Right now this is all the samples in the cell-gene dataset.
```{r}
samples = c("Uninfected", "6hr-1", "8hr-2", "8hr-2repeat", "10hr-2")

cellfilter <- row.names(subset(pData(all_cells), Sample %in% samples))
cells <- all_cells[,cellfilter]

cat("Filtered all", nrow(pData(all_cells)), "cells down to the", nrow(pData(cells)), 
      "in samples of interest.")
```

Compute the number of total mRNAs, cellular mRNAs, influenza mRNAs, and fraction of influenza mRNAs per cell.
```{r}
pData(cells)$total_mRNAs <- Matrix::colSums(exprs(cells))

flugenes <- c("fluPB2", "fluPB1", "fluPA", "fluHA", "fluNP", "fluNA", "fluM", "fluNS")
flugenes_id <- row.names(subset(fData(cells), gene_short_name %in% flugenes))
pData(cells)$flu_mRNAs <- Matrix::colSums(exprs(cells[flugenes_id,]))

pData(cells)$frac_flu_mRNAs <- pData(cells)$flu_mRNAs / pData(cells)$total_mRNAs

cellgenes_id <- row.names(subset(fData(cells), !(gene_short_name %in% flugenes)))
pData(cells)$cellular_mRNAs <- Matrix::colSums(exprs(cells[cellgenes_id,]))
```

Plot the number of influenza mRNAs versus the number of cellular RNAs.
The blue curves show the overall density of the number of cellular mRNAs.
Note that different samples were sequenced to somewhat different depths.
```{r, fig.height=4, fig.width=6.5}
scatterplot <- ggplot(pData(cells), aes(cellular_mRNAs, flu_mRNAs)) + geom_point()

# next line uses eval substitute to get value of maxy as here: https://stackoverflow.com/a/10659563
densityplot <- eval(substitute(
    {stat_density(aes(x=cellular_mRNAs, y=maxy*(..scaled..)), geom="line", color='blue')}, 
    list(maxy=layer_scales(scatterplot)$y$range$range[2])))

scatterplot + densityplot + theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + 
    facet_wrap(~Sample) + xlab("cellular mRNAs per cell") + ylab("influenza mRNAs per cell")
```

The `Monocle` vignette suggests filtering cells with far more / less mRNAs than most cells.
I have **not** done that because it filters out a lot of the influenza-infected cells.
This fact is made more clear by the next plot, which shows that cells with >1% of their mRNAs from influenza tend to have more total mRNAs than other cells.
I can think of several possible explanations:

  1. Influenza preferentially infects bigger cells.

  2. Influenza infection leads cells to produce more mRNAs.

  2. Influenza infection preferentially causes "doublets"" due to cell clumping, perhaps mediated by hemagglutinin-based cross-linking of cells.

In any case, it seems safer **not** to exclude the influenza-infected cells as we'd then be preferentially excluded cells with high viral burden.
```{r, fig.height=2.75, fig.width=5.5}
pData(cells)$frac_flu_0.01 <- ifelse(pData(cells)$frac_flu_mRNAs > 0.01, 
    "more than 1%", "less than 1%")

ggplot(pData(cells[, row.names(subset(pData(cells), Sample %in% c("8hr-2", "10hr-2")))]), 
        aes(total_mRNAs, color=frac_flu_0.01, fill=frac_flu_0.01)) +
    geom_density(alpha=0.2) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    scale_y_continuous(label=NULL, name="density of cells", breaks=NULL, limits=c(0, NA)) + 
    scale_x_continuous(limits=c(0, NA)) +
    xlab("total mRNAs per cell") + 
    labs(color="mRNAs from\ninfluenza", fill="mRNAs from\ninfluenza") + 
    facet_wrap(~Sample)
```

Here are some cumulative distribution plots showing the fraction of mRNAs derived from influenza across cells for each sample.
For the 6 and 8 hour timepoints, less than a quarter of cells have any appreciable influenza mRNAs.
For the 10 hour timepoint, about a quarter the cells have at least a tiny fraction of influenza mRNAs, but only about 10% have more than 0.1% of mRNAs from influenza.
```{r, fig.width=5, fig.height=2.75}
# some trickery around x minimum to avoid taking log of 0 while plotting x on log scale
minx <- 1e-4 # minimum value on y-axis
pData(cells)$log_frac_flu_plus <- (function(x){ifelse(x < minx, log10(minx), log10(x))}
    )(pData(cells)$frac_flu_mRNAs)

ggplot(pData(cells), aes(log_frac_flu_plus, color=Sample)) + 
    stat_ecdf() + 
    scale_x_continuous(labels=function(x){fancy_scientific(10**x)},
        limits=c(log10(minx), 0), expand=c(0, 0), name="fraction mRNAs from influenza") +
    scale_y_continuous(limits=c(0, 1), name="fraction cells with > this\nmany influenza mRNAs") 
```