### 2017-08-10 paper/figures/Pseudovirus_flow_cytometry/Analysis.Rmd
Figures and data added to paper

---
title: "Flow_Analysis"
author: "Alistair Russell"
date: "7/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup}
# pipe operators
suppressPackageStartupMessages(library(magrittr))
# for manipulating factors
suppressPackageStartupMessages(library(forcats))
# tab data processing
suppressPackageStartupMessages(library(tidyverse))
# supress warnings
options(warn = -1)
# for low-level manipulation of ggplot
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(ggforce))
library(grid)
#loading flow cytometry data
suppressPackageStartupMessages(library(flowCore))
```

## Including Plots

You can also embed plots, for example:

```{r load data}
# look for files ending in .fcs
fcsfiles <- 'Data/HA' %>% 
    list.files(full.names = T, pattern = '.fcs$')

flowdata  <- fcsfiles %>% 
    lapply(. %>%
           # read FCS file without any transformations
           flowCore::read.FCS(transformation = F, alter.names = T) %>% 
           # extract counts alone
           flowCore::exprs() %>% 
           # convert to dplyr tibble
           as_tibble() 
          )

names(flowdata) <- fcsfiles %>% 
    # extract file name without .fcs extension and parent folders
    lapply(. %>% 
           stringr::str_match('([^/]+).fcs$') %>% 
           extract2(2))

# join all samples by 'file'
flowdata %<>% bind_rows(.id = 'file')

print(flowdata)
write.table(flowdata, 'coinfection.tsv', sep='\t', row.names=FALSE)
```