### 2017-09-13 paper/paper.tex
re-draft abstract and work on beginning of Discussion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ELIFE ARTICLE TEMPLATE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE 
\documentclass[9pt,lineno]{elife}
% Use the onehalfspacing option for 1.5 line spacing
% Use the doublespacing option for 2.0 line spacing
% Please note that these options may affect formatting.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\DeclareSIUnit\Molar{M}
%\usepackage{indentfirst} %I prefer to have indents following section headings, I find it to be more consistent in style throughout document
\usepackage{natbib}
\usepackage{float}

\newfloat{suppfile}{thp}{lofsupfile}
%\renewcommand{\thesuppfile}{Supplementary file \arabic{suppfile}}
\floatname{suppfile}{Supplementary file}

\renewcommand{\floatpagefraction}{1}

\newcommand{\jdbcomment}[1]{\emph{\color{red} [#1]}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Extreme heterogeneity of influenza virus transcriptional load in single cells}

\author[1]{Alistair B. Russell}
\author[2]{Cole Trapnell}
\author[1,2*]{Jesse D. Bloom}
\affil[1]{Basic Sciences Division and Computational Biology Program, Fred Hutchinson Cancer Research Center, Seattle, United States}
\affil[2]{Department of Genome Sciences, University of Washington, Seattle, United States}
\corr{jbloom@fredhutch.org}{}

% \presentadd[\authfn{5}]{eLife Sciences editorial Office, eLife Sciences, Cambridge, United Kingdom}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

\begin{abstract}
Viral infection can cause large changes in a cell's transcriptome, but these changes have mostly been studied by bulk measurements on many cells.
We use single-cell mRNA sequencing to examine the consequences of influenza virus infection in individual cells.
We find extremely wide cell-to-cell variation in viral gene expression -- viral transcripts compose no more than a few percent of total mRNA in most infected cells, but a few cells derive over half their mRNA from virus.
Many infected cells fail to express at least one viral gene, and this gene absence explains some of the variation in viral transcriptional load.
Activation of innate immune pathways is rare, even in cells that fail to express influenza virus's major interferon antagonist.
We identify some cellular genes that co-vary in abundance with the amount of viral mRNA.
Overall, our results highlight the extensive cell-to-cell heterogeneity in the transcriptional consequences of viral infection.
\end{abstract}

\section{Introduction}
Viruses can cause massive and rapid changes in a cell's transcriptome as they churn out viral mRNAs and hijack cellular machinery.
For instance, cells infected with influenza virus at high multiplicity of infection (MOI) express an \emph{average} of 50,000 to 100,000 viral mRNAs per cell, corresponding to 5 to 25\% of all cellular mRNA~\citep{Hatada:1989vz}.
In addition, infection can trigger innate-immune sensors that induce the expression of cellular anti-viral genes~\citep{Killip:2015dw,Iwasaki:2014dw,Crotta:2013ef}.
This anti-viral response is another prominent transcriptional signature of high-MOI influenza virus infection in bulk cells~\citep{Geiss:2002a}.

However, initiation of an actual influenza infection typically involves just a few virions infecting a few cells~\citep{varble2014influenza,mccrone2017evolutionary}. 
The dynamics of viral infection in these individual cells may not mirror bulk measurements made on many cells infected at high MOI.
Over 70 years ago, \citet{Delbruck:1945wo} showed that there was a $\sim$100-fold range in the number of progeny virions produced per cell by clonal bacteria infected with clonal bacteriophage.
Subsequent work has shown similar heterogeneity in the number of progeny virions per cell during infection with other viruses~\citep{Zhu:2009ea,Schulte:2014br,combe2015single,Akpinar:2016gt}, including influenza virus~\citep{Heldt:2015iz}.

In the case of influenza virus infection, targeted single-cell measurements of specific proteins or RNAs have shed light on some factors that contribute to cell-to-cell heterogeneity. 
The influenza virus genome consists of eight negative-sense RNA segments, and many infected cells fail to express one more of these RNAs~\citep{Heldt:2015iz,Dou:2017cp} or their encoded proteins~\citep{Brooke:2013kb}.
In addition, the activation of innate-immune responses is inherently stochastic~\citep{Shalek:2013ej,Shalek:2014ey,bhushal2017cell}, and only some influenza-infected cells express anti-viral interferon genes~\citep{PerezCidoncha:2014jr,Killip:2017ef}.
However, the extent of cell-to-cell variation in these and other host and viral factors remains unclear, as does the association among them in individual infected cells.

Here we use single-cell mRNA sequencing to quantify the levels of all cellular and viral mRNAs in cells infected with influenza virus at low MOI.
We find extremely large variation in the amount of viral mRNA expressed in individual cells.
Both co-infection and activation of innate-immune pathways are rare in our low-MOI infections, and do not appear to be the major drivers of cell-to-cell heterogeneity in viral transcriptional load.
Individual infected cells often fail to express specific viral genes, and such gene absence explains some but certainly not all of the cell-to-cell heterogeneity.
A variety of cellular genes, including ones involved in the oxidative-stress response, co-vary with viral transcriptional load.
Overall, our work demonstrates remarkable heterogeneity in the transcriptional outcome of influenza virus infection among nominally identical cells infected with a relatively pure population of virions under controlled conditions.

\section{Results}

\subsection{Strategy to measure mRNA in single virus-infected cells.}
We performed single-cell mRNA sequencing using a droplet-based system that physically isolates individual cells prior to reverse transcription~\citep{zheng2017massively,Macosko:2015ht,Klein:2015ki}.
Each droplet contains primers with a unique \emph{cell barcode} that tags all mRNAs from that droplet during reverse-transcription.
Each primer also contains a \emph{unique molecular identifier (UMI)} that is appended to each mRNA molecule during reverse transcription.
The 3' ends of the mRNAs are sequenced and mapped to the human and influenza virus transcriptomes to determine transcript identities.
This information is combined with that provided by the UMIs and cell barcodes to quantify the number of molecules of each mRNA species that have been captured for each cell.

Infected cells will express viral as well as cellular mRNAs -- however the cell barcodes and UMIs cannot distinguish whether a cell was initially infected by one or multiple viral particles.
We therefore engineered an influenza virus (strain A/WSN/1933) that additionally carried \emph{viral barcodes} consisting of synonymous mutations near the 3' end of each transcript (Figure~\ref{fig:workflow}A).
Critically, these synonymous mutations did not greatly impact viral growth kinetics (Figure~\ref{fig:workflow}B).
We infected A549 human lung carcinoma cells with an equal mix of the wild-type and synonymously barcoded viruses.
Cells infected by a single virion will exclusively express mRNAs from either wild-type or synonymously barcoded virus, whereas cells that are co-infected with multiple virions will often express mRNAs from both the wild-type and synonymously barcoded viruses (Figure~\ref{fig:workflow}C).
\begin{figure}
\centerline{\includegraphics[width=0.8\linewidth]{figures/Workflow/workflow.pdf}}
\caption{\label{fig:workflow} Experimental design.
{\bf (A)}  
We engineered a virus that carried two synonymous mutations near the 3' end of each mRNA.
At top are the specific mutations for PB2.
At bottom are locations of the synonymous mutations relative to the typical distribution of read depth for our 3'-end sequencing.
{\bf (B)} 
The wild-type and synonymously barcoded viruses transcribe their genes with similar kinetics. 
Shown is the abundance of the viral hemagglutinin (HA) transcript relative to the cellular housekeeping gene L32 as assessed by qPCR in A549 cells infected at an MOI of 0.5 (as determined on MDCK-SIAT1 cells).
Error bars $\pm$ S.D., n=3.
{\bf (C)}  
For the single-cell mRNA sequencing, A549 cells were infected with an equal mixture of mutant and wild-type virus. 
Immediately prior to collection, cells were physically separated into droplets and cDNA libraries were generated containing the indicated barcodes. 
The libraries were deep sequenced, and the data processed to create a matrix that gives the number of molecules of each transcript observed in each cell.
Infected cells were further annotated by whether their viral mRNAs derived from wild-type virus, synonymously barcoded virus, or both.
}
\figdata{Sequences of wild-type and barcoded viruses are in \texttt{viralsequences.fasta}.\label{figdata:viralsequences}}
\end{figure}

We took care to generate stocks of virus that were relatively ``pure'' of defective particles.
Stocks of viruses typically contain an array of biologically active viral particles, some of which are defective for replication owing to mutations or deletions in essential viral genes~\citep{von1954incomplete,huang1970defective,Brooke:2014ch,Fonville:2015cg,Lauring:2010if,Dimmock:2014dk}.
These defective particles become prevalent when a virus is grown at high MOI, where complementation permits the growth of otherwise deleterious genotypes.
To minimize the levels of defective particles, we propagated our viral stocks at low MOI for a relatively brief period of time~\citep{Xue:2016bw}.
We validated that our stocks exhibited greater purity of infectious particles than a stock propagated at high MOI by verifying that they had a higher ratio of infectious particles to virion RNA (Figure~\ref{fig:viruspopulations}A) and to particles capable of inducing expression of a single viral protein (Figure~\ref{fig:viruspopulations}B).
In addition, viral stocks with many defective particles are more immunostimulatory~\citep{Tapia:2013kf,Lopez:2014en}.
We confirmed that our viral stocks induced much less interferon than a stock propagated at higher MOI (Figure~\ref{fig:viruspopulations}C).

\begin{figure}
\centerline{\includegraphics[width=0.7\linewidth]{figures/Validating_barcode_virus/validating_populations_D02.pdf}}
\caption{\label{fig:viruspopulations} The viral stocks in our experiments are relatively pure of defective particles. 
{\bf (A)}
Our viral stocks have a higher ratio of infectious particles to HA virion RNA compared to a high-defective stock propagated at high MOI.
HA viral RNA was quantified by qPCR on virions. 
Error bars $\pm$ S.D., n=6 (qPCR replicates). 
{\bf (B)} 
Our viral stocks have a higher ratio of infectious particles to particles capable of expressing the viral HA protein.
A549 cells were infected with all three viral stocks at an MOI of 0.1, and the percentage of cells expressing HA protein at 9 hours post-infection was quantified by antibody staining and flow cytometry.
{\bf (C)} 
Our viral stocks are less immunostimulatory than virus propagated at high MOI. 
Measurements of \textit{IFNB1} transcript by qPCR normalized to the housekeeping gene L32 in A549 cells at 10 hours post infection at an MOI of 0.5.
Error bars $\pm$ S.D., n=3.
Note that the MOIs were calculated by TCID50 on MDCK-SIAT1 cells, whereas the experiments in this figure involved infection of A549 cells.
}
\figsupp[Full flow cytometry data for panel B . \label{figsupp:HAstain}]{
Full flow cytometry date for  Figure~\ref{fig:viruspopulations}B. 
A549 cells were infected at an MOI of 0.1 as calculated by TCID50 on MDCK-SIAT1 cells. 
{\bf (A)} Uninfected gating control.
{\bf (B)} Cells infected with the wild-type virus stock used in our experiments.
{\bf (C)} Cells infected with synonymously barcoded virus stock used in our experiments.
{\bf (D)} Cells infected with a stock of wild-type virus propagated at a high MOI, and therefore enriched in defective particles.
}{\includegraphics[width=0.7\linewidth]{figures/Validating_barcode_virus/HA_stain_supplement.png}}\end{figure}

\subsection{Single cells show an extremely wide range of expression of viral mRNA.}
We infected A549 cells at low MOI with a mixture of the wild-type and synonymously barcoded viruses, and collected cells for sequencing at 6, 8, and 10 hours post-infection, including a replicate of the 8-hour timepoint.
For all samples except for the replicate 8-hour timepoint, we replaced the infection inoculum with fresh media after one hour, thereby ensuring that most infection was initiated during a narrow time window.
We recovered between 3,000 and 4,000 cells for each sample (Figure~\ref{fig:cells}A). 
As expected for a low-MOI infection, most cells expressed little or no viral mRNA (Figure~\ref{fig:cells}B).
Also as expected, the amount of viral mRNA per cell among infected cells increased over time (Figure~\ref{fig:cells}B).
But what was most notable was how widely the number of viral mRNA molecules varied among infected cells.
While the fraction of mRNA derived from virus was $<$0.1\% for most cells, viral mRNA constituted half the transcriptome in a few cells at 8 and 10 hours (Figure~\ref{fig:cells}C).

\begin{figure}
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_cell_mRNA_summary.pdf}}
\caption{\label{fig:cells}
There is a very wide distribution in the number of viral mRNAs detected per cell.
{\bf (A)} 
Number of cells sequenced for each sample.
{\bf (B)} 
Cumulative distribution of the fraction of mRNA derived from influenza virus for each sample.
For each point on the x-axis, the y-axis indicates the fraction of cells with at least that much of their mRNA from virus.
{\bf (C)} 
The number of cellular and viral mRNAs for each cell is plotted as a point.
The blue lines show the overall distribution of the number of cellular mRNAs per cell.
Cells outside the red lines were considered outliers (possibly derived from two cells per droplet), and were excluded from subsequent analyses.
}
\end{figure}

A complicating factor is that uninfected cells could have small amounts of viral mRNA due to leakage of transcripts from lysed cells.
It is therefore important to establish a threshold for identifying truly infected cells.
We can do this by taking advantage of the fact that roughly half the infecting virions bear synonymous barcodes.
Reads derived from lysed cells will be drawn from both wild-type and synonymously barcoded viral transcripts.
However, most cells are infected by at most one virion, and so the reads from truly infected cells will usually derive almost entirely from one of the two viral variants. 
Figure~\ref{fig:viralbarcodes}A shows the fraction of viral reads in individual cells from each viral variant, and Figure~\ref{fig:viralbarcodes}B indicates the fraction of viral reads from the most abundant variant in that cell.
Most cells with large amounts of viral mRNA have viral transcripts exclusively derived from one viral variant -- indicating non-random partitioning as expected from viral infection.
However, cells with a small amount of viral mRNA often have viral transcripts from both variants, as expected from the random partitioning associated with simple mRNA leakage.
Finally, a few cells with large amounts of viral mRNA have viral transcripts from both variants, reflecting co-infection.

We determined the threshold amount of viral mRNA per cell at which the barcode partitioning clearly resulted from infection rather than leakage (Figure~\ref{fig:viralbarcodes}C), and used this threshold to annotate cells that we were confident were truly infected.
We also annotated as co-infected cells above this threshold that had mRNA from both viral variants.
Figure~\ref{fig:viralbarcodes}D shows the number of cells annotated as infected and co-infected for each sample -- these cells are just a small fraction of the number of cells with any viral read.
These annotation thresholds are conservative, and will miss some true low-level infections, as well as any co-infections with the same viral variant.
However, it is important that the analyses below are restricted to cells that are truly infected with virus, so we accepted the loss of some low-level infections in order to avoid false positives.
Because most cells are not infected, we subsampled the uninfected cells to the numbers shown in Figure~\ref{fig:viralbarcodes}D to balance the proportions of infected and uninfected cells for all subsequent analyses.

Strikingly, the extreme variation in the number of viral transcripts per cell remains even after we apply these rigorous criteria for annotating infected cells (Figure~\ref{fig:viralbarcodes}E). 
The fraction of viral mRNA per infected cell follows a roughly exponential distribution, with many cells having few viral transcripts and a few cells having many.
At 6 and 8 hours $<$10\% of infected cells are responsible for over half the viral transcripts, while at 10 hours $\approx$15\% of infected cells produce over half the viral transcripts (Figure~\ref{fig:viralbarcodes}-Figure~supplement~\ref{figsupp:cumulfracflu}).
Notably, co-infected cells do not express dramatically more viral mRNA than other cells (Figure~\ref{fig:viralbarcodes}), suggesting that the initial infectious dose is not the major driver of cell-to-cell variation at least in our low-MOI infections.

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_frac_flu_summary.pdf}}
\caption{\label{fig:viralbarcodes}
Synonymous barcodes on the viral mRNAs distinguish true infections from cells that contain viral mRNAs derived from leakage of lysed cells.
{\bf (A)}
Cells with at least two viral mRNAs for which the barcode could be called, arranged in order of increasing influenza transcript counts.
Bar heights denote the number viral mRNAs on a log\textsubscript{10} scale, bar coloring is linearly proportional to the fractions of viral mRNAs derived from wild-type and synonymously barcoded virus.
{\bf (B)}
Same as (A), but each bar is colored according to the relative fraction of the more common (major) and less common (minor) virus variant.
At low levels of viral mRNA there is often a roughly equal mix, suggesting contamination with viral mRNAs leaked from lysed cells.
At higher levels of viral mRNA, cells generally have only one viral variant, suggesting infection initiated by a single particle.
A few cells are also obviously co-infected with both viral variants.
{\bf (C)}
We determined a cutoff for calling ``true'' infections by fitting a curve to the mean barcode purity of all cells with greater than a given fraction of their mRNA derived from virus.
We called the cutoff at the point at which purity stops increasing with the fraction of viral mRNA.
{\bf (D)}
The number of cells identified as infected and co-infected for each sample, as well as the number of cells with any viral read.
For all subsequent analyses, we subsampled the number of uninfected cells per sample to the greater of 50 or the number of infected cells.
{\bf (E)} 
Distribution of the fraction of mRNA per cell derived from virus for both infected and co-infected cells.
}
\figsupp[Number of viral barcodes called.\label{figsupp:barcodescalled}]{
The number of viral barcodes called for each sample and gene segment. 
Viral transcripts are classified as \emph{syn} if they mapped to a synonymously barcoded influenza transcript, \emph{wt} if they mapped to a wild-type influenza transcript, \emph{invalid} if multiple reads for the same UMI differed on the status of the viral barcode, and as \emph{uncalled} if none of the reads for that UMI overlapped the region of the viral transcript containing the viral barcode.
}{\includegraphics[width=\linewidth]{figures/synbarcodes_umistats.jpg}}
\figsupp[Fraction of total viral mRNA derived from a given fraction of infected cells.\label{figsupp:cumulfracflu}]{
The total fraction of all viral mRNA among infected cells that is attributable to a given fraction of these cells.
For instance, the plot for the 8-hour sample shows that roughly 50\% of all viral mRNA is derived from 10\% of the infected cells.
}{\includegraphics[width=\linewidth]{figures/p_cumul_flu.pdf}}
\end{figure}

\subsection{Absence of viral genes partially explains cell-to-cell variability in viral transcript abundance.}
The influenza genome is segmented, and cells can fail to express a viral mRNA if the encoding gene segment is not packaged in the infecting virion or fails to initiate transcription after infection.
Indeed, several groups have reported that the majority of infected cells fail to express at least one viral gene~\citep{Brooke:2013kb, Heldt:2015iz,Dou:2017cp}. 
We wondered if the absence of specific viral genes might be associated with reduced amounts of viral mRNA within single infected cells.
In particular, transcription of influenza virus mRNAs is performed by the viral ribonucleoprotein (RNP) complex, which consists of the three proteins that encode the tripartite polymerase (PB2, PB1, and PA) as well as nucleoprotein (NP)~\citep{huang1990determination}.
Each viral gene segment is associated with one RNP in incoming infecting virions, but secondary transcription by newly synthesized RNPs requires the presence of the viral genes encoding each of the four RNP proteins~\citep{Vreede:2004ip,eisfeld2015centre}.
This secondary transcription is a major source of viral mRNAs, as evidenced by the fact that blocking synthesis of the RNP proteins reduces the amount of viral mRNA by several orders of magnitude in bulk cells (Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:cyclohexamide}).

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_flu_burden_flu_gene_merge.pdf}}
\caption{\label{fig:fluburdenbyflugene}
The absence of viral genes explains some of the variability in the amount of viral mRNA per cell.
{\bf (A)} 
Fraction of mRNA in each infected cell derived from virus as a function of the normalized expression of each viral gene in that cell, taken over all time points.
Cells with high viral burden always express all RNP genes, but some cells with high viral burden lack each of the other genes.
{\bf (B)}
Box plots showing the per-cell viral burden among cells with $>$0.5\% of their mRNA from virus, binned by whether or not the cells express each gene.
A Wilcoxon signed-rank test was used to test the null hypothesis that absence of each gene does not affect viral burden: **** = $P < 10^{-4}$, *** = $P < 10^{-3}$,  * = $P < 0.05$, ns = not significant.
Similar results are obtained if we examine only the 10-hour timepoint (Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:10hrfluburdenbyflugene}).
Top and bottom of box bound the first and third quartiles; the line in the box is at the median.
Whiskers extend to the highest or lowest point within1.5$\times$ the interquartile range, and outliers are shown as circles.
{\bf (C)}
The fraction of that express each of the four other genes among cells that express all RNP genes, as well as the fraction that express \emph{all} four of the other genes.
The fraction that express all four genes is well predicted by simply multiplying the frequencies of cells that express each gene individually, indicating that gene absence is approximately independent across these genes.
}
\figsupp[Secondary transcription from newly synthesized RNPs is a major source of viral mRNA during bulk infections.\label{figsupp:cyclohexamide}]{
A549 cells were infected at an MOI of 0.2 as calculated on MDCK-SIAT1 cells in either the presence or absence of the protein-translation inhibitor cyclohexamide, and viral mRNA was quantified at 8 hours post-infection by qPCR.
The cyclohexamide prevents translation of new PB2, PB1, PA, and NP protein, and so prevents the formation of the new RNPs needed for secondary transcription.
The bars show the relative amount of HA and PB2 mRNA in the absence versus the presence of cyclohexamide. Error $\pm$ S.D. n=3. 
}{\includegraphics[width=0.3\linewidth]{figures/Primary_transcription_measurements/Cyclohexamide.pdf}}
\figsupp[Like panel (B) but for the 10-hr sample only.\label{figsupp:10hrfluburdenbyflugene}]{
The absence of viral RNP genes but \emph{not} non-RNP genes remains significantly associated with reduced viral burden when we examine only the 10-hr sample, which is the single time point with the most data points.
The difference for NP is no longer statistically significant due to low counts of infected cells lacking NP, but the trend remains.
}{\includegraphics[width=\linewidth]{figures/p_10hr_flu_burden_flu_gene_test}}
\figdata{The numerical data for panel (C) are in \texttt{p\_missing\_genes.csv}.}
\end{figure}

We examined the total amount of viral mRNA versus the expression of each viral gene (Figure~\ref{fig:fluburdenbyflugene}A). 
Cells that lack an RNP gene never derive more than a few percent of their mRNAs from virus, confirming the expected result that all four RNP genes are essential for high levels of viral transcription.
However, we observe cells that lack each of the other non-RNP genes but still derive $\approx$40\% of their mRNAs from virus, suggesting that none of the other genes are important for high levels of viral transcription.
These results are statistically supported by Figure~\ref{fig:fluburdenbyflugene}B, which shows that absence of any RNP gene but \emph{not} any other viral gene is associated with reduced amounts of viral mRNA. 
However, gene absence clearly does not explain all of the variability in viral gene expression, since even cells expressing all viral genes exhibit a very wide distribution in the amount of viral mRNA that they express. 
The rest of this variability must be due to additional mechanisms such as differences in host-cell state, viral mutations not detectable by our 3'-end sequencing approach, or inherent stochasticity.

We also quantified the fraction of infected cells that completely failed to express a given gene.
We limited this analysis to examining the presence / absence of the non-RNP genes in cells expressing all four RNP genes, since we might fail to detect viral transcripts that are actually present at low levels in RNP-deficient cells due to the lower viral burden in these cells.
At the 8- and 10-hour time points, between 5\% and 17\% of cells fail to express any one of the four non-RNP genes (Figure~\ref{fig:fluburdenbyflugene}C).
The absence of a given gene appears to be an independent event, as the probability of observing all four non-RNP genes in a cell is well predicted by simply multiplying the probabilities of observing each gene individually (Figure~\ref{fig:fluburdenbyflugene}C). 
If we extrapolate the frequencies at which cells lack non-RNP genes to the RNP genes, then we would predict that 40-50\% of infected cells express mRNAs from all eight genes.
This estimate of the frequency at which infected cells express mRNAs from all eight gene segments is slightly higher than previous estimates of 13\%~\citep{Brooke:2013kb} and 20\%~\citep{Dou:2017cp}.
At least one difference is that \citet{Brooke:2013kb} stained for proteins whereas we examined the expression of mRNAs -- it is likely that some cells contain mutated viral genes that fail to produce stable protein even when mRNA is expressed.

\subsection{The relative amounts of different viral mRNAs are more consistent across cells.}
The results above show that the amount of viral mRNA in infected cells varies over several orders of magnitude.
Does the relative expression of viral genes exhibit similar cell-to-cell variability?
To address this question, we focused on cells that derived $>$5\% of their mRNA from virus, since estimates of relative viral gene expression will be less noisy in cells with more viral mRNAs.

\begin{figure}
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_flu_expr_aledit.pdf}}
\caption{\label{fig:fluexpr}
Relative expression of influenza virus genes in highly infected cells (>5\% of total mRNA from virus).
{\bf (A)} 
The fraction of viral mRNA from each viral gene for each cell. 
{\bf (B)}
Box plots showing the distribution of the fraction of viral mRNA per cell from each viral gene.
The black lines at the notches are the medians, and the tops and bottoms of boxes indicate the first and third quartiles.
Whiskers extend to the highest or lowest data point observed within 1.5x the interquartile range, outliers shown as circles.
Notches extend 1.58x the interquartile range divided by the square root of the number of observations. 
{\bf (C)}, {\bf (D)} 
The same plots, but only including cells for which we observed at least one molecule of each viral gene.
}
\figdata{The raw data are in \texttt{p\_flu\_expr.csv}.}
\end{figure}

In contrast to the extreme variability in the total viral mRNA per cell, the fraction of this mRNA derived from each gene is much more consistent across cells (Figure~\ref{fig:fluexpr}A).
Total viral mRNA varies by orders of magnitude, but the fraction from any given viral gene is fairly tightly clustered around the median value for all cells (Figure~\ref{fig:fluexpr}B).
The relative levels of each viral mRNA in our cells are similar to prior bulk measurements made by Northern blots~\citep{Hatada:1989vz}, which also found an expression hierarchy of M $>$ NS $\gg$ NP $>$ NA $>$ HA $\gg$ PB2 $\sim$ PB1 $\sim$ PA.
The cell-to-cell consistency in the relative expression of different viral genes is even tighter if we limit the analysis only to cells that express all eight viral genes (Figure~\ref{fig:fluexpr}C,D).
Therefore, with the exception of complete gene absence, the factors that drive the dramatic cell-to-cell variability in the amount of viral mRNA have roughly similar effects on all viral genes in a given cell.
This finding is consistent with prior work showing positive correlations among the abundance of several viral genes in individual cells~\citep{Heldt:2015iz}.

\subsection{Co-infection can provide infected cells with the full complement of viral genes.}
Our sequencing enables us to identify the rare cells that were co-infected with both wild-type and synonymously barcoded viral variants.
Overall, we captured 10 such co-infected cells that had $>$5\% of their mRNA derived from virus (Figure~\ref{fig:coexpression}).
Seven of these 10 cells expressed all eight viral genes.
Remarkably, the majority (4 of 7) of these cells would \emph{not} have expressed all the viral genes in the absence of co-infection, since they have at least one gene exclusively derived from each viral variant.
For instance, the cell with 11.2\% of its mRNA from virus in the upper right of Figure~\ref{fig:coexpression} expresses M only from the wildtype viral variant, and NP and HA only from the synonymously barcoded variant.
Our data therefore provide direct single-cell support for the idea that co-infection can rescue missing viral genes~\citep{Brooke:2013kb,brooke2014influenza,Fonville:2015cg,aguilera2017plaques}.

\begin{figure}
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_coinfection.pdf}}
\caption{\label{fig:coexpression}
The abundance of each viral transcript in cells that are co-infected with the two viral variants and have $>$5\% of their mRNA derived from virus.
The bars show the logarithms of the numbers of each viral mRNA detected, and are colored in linear proportion to the fraction of that mRNAs derived from wild-type or synonymously barcoded virus.
}
\figsupp[Co-infected cells express roughly equal amounts of a gene from each infecting viral variant. \label{figsupp:flowcyto}]{
{\bf(A)} Cells were co-infected with a mix of wild-type virus and virus in which the HA gene was replaced by GFP flanked by the terminal regions of the HA gene segment. 
At 10 hours post-infection, cells were analyzed by flow cytometry for HA and eGFP expression.
{\bf(B)} The expression of HA and GFP are correlated in co-infected cells. 
Shown are the quantile-normalized HA and eGFP signals for double-positive cells. 
Cells are colored by density, using a Gaussian kernel density estimate. 
{\bf(C)},{\bf(D)},{\bf(E)} Gating controls, single infection with eGFP virus, single infection with wild-type virus, and uninfected cells, respectively. 
}{\includegraphics[width=\linewidth]{figures/Pseudovirus_flow_cytometry/coinfFlow_D02.png}}
\figdata{The raw data plotted in this figure are in \texttt{p\_co-infection.csv}.}
\figdata{The sequence of the HA viral RNA carrying the GFP gene is in \texttt{HAflank-eGFP.fasta}.\label{figdata:HAflank}}
\end{figure} 
Another observation from Figure~\ref{fig:coexpression} is that co-infected cells usually express roughly equal amounts of transcripts from each of the two viral variants.
This observation is consistent with the finding by \citet{Dou:2017cp} and \citet{Huang:2008gy} that the temporal window for co-infection is short -- if both viral variants infect a cell at about the same time, then neither will have a headstart and so each will have a roughly equal opportunity to transcribe its genes.

To support this idea with a larger dataset albeit at lower resolution, we generated a virus in which the HA coding sequence was replaced by GFP.
We then co-infected cells with a mix of wildtype and $\Delta$HA-GFP virus and used flow cytometry to score cells for the presence of HA only (infection by wildtype virus), GFP only (infection by $\Delta$HA-GFP virus), or both (co-infection) as shown in Figure~\ref{fig:coexpression}~Figure~supplement~\ref{figsupp:flowcyto}.
As in our single-cell sequencing data, we found that expression of HA and GFP were highly correlated, indicating that co-infected cells typically expressed roughly equal amounts of transcript from each viral variant.

\subsection{Activation of the interferon response is rare in single infected cells.}
Because our sequencing captured all polyadenylated transcripts, we can examine whether there are prominent changes in the host-cell transcriptome in sub-populations of infected cells.
Influenza virus infection can trigger innate-immune sensors that lead to the transcriptional induction of type I and III interferons, and subsequently of anti-viral interferon-stimulated genes~\citep{Killip:2015dw,Iwasaki:2014dw,Crotta:2013ef}.
%In bulk cells, this interferon response is one of the most prominent transcriptional signatures of influenza-virus infection~\citep{Geiss:2002a}.
However, activation of the interferon response is stochastic and bi-modal at the level of single cells~\citep{Chen:2010cr,Shalek:2013ej,Shalek:2014ey,PerezCidoncha:2014jr,bhushal2017cell}.
We therefore hypothesized that we might see two sub-populations of infected cells: one in which the interferon response inhibited viral transcription, and another in which the virus was able to express high levels of its mRNA by evading or blocking this response.

\begin{figure}
\centerline{\includegraphics[width=0.8\linewidth]{figures/p_small_tsne_merge.pdf}}
\caption{\label{fig:tsne}
A t-SNE plot created by semi-supervised clustering using genes that co-vary with viral infection status.
Each point is a single cell, and each panel shows an identical layout but colors the cells according to a different property.
{\bf (A), (B)}
Cells colored by the fraction of their mRNA derived from virus.
{\bf (C)}
Cells colored by the experimental sample.
{\bf (D), (E)}
Cells colored by the number of detected transcripts from type I and III interferons (IFN).
Only one cell has detectable interferon expression.
{\bf (E)}
Cells colored by the expression of the interferon-stimulated gene IFIT1.
{\bf (F)}
Cells colored by whether they express the viral NS gene.
The one interferon-positive cell is lacking NS, but many interferon-negative cells also lack NS.
}

\end{figure}

To examine whether there were distinct sub-populations of virus-infected cells, we used a semi-supervised t-SNE approach~\citep{VanderMaaten:2008tm} to cluster cells by genes that co-varied with viral infection status.
As shown in Figure~\ref{fig:tsne}A,B, this approach effectively grouped cells by the amount of viral mRNA that they expressed.
Sample-to-sample variation was regressed away during the clustering, as cells did not obviously group by time-point, with expected exception that the uninfected and 6-hour samples had few cells in the region of the plot corresponding to large amounts of viral mRNA (Figure~\ref{fig:tsne}C).

But to our surprise, we did not see a prominent clustering of infected cells into sub-populations as expected if the interferon response was strongly activated in some cells.
To investigate further, we annotated each cell by the total number of type I and III interferon transcripts detected.
Remarkably, only a single cell expressed detectable interferon (Figure~\ref{fig:tsne}D).
We also examined interferon-stimulated genes, which are induced by autocrine and paracrine interferon signaling.
Figure~\ref{fig:tsne}E shows expression of one such gene, IFIT1~\citep{Fensterl:2011fp}.
As with interferon itself, expression of IFIT1 was rare and most prominent in the single interferon-positive cell, presumably due to the higher efficiency of autocrine versus paracrine signaling.
Notably, interferon and interferon-stimulated genes were also relatively ineffective at blocking viral transcription in the single cell in which they were potently induced, since $>$10\% of the mRNA in this cell was derived from virus (Figure~\ref{fig:tsne}A,B,D,E).

We posited that the paucity of interferon induction might be due to the activity of influenza virus's major interferon antagonist, the NS1 protein~\citep{garcia1998influenza,hale2008multifunctional}.
We therefore identified cells that expressed substantial amounts of viral mRNA but lacked the NS gene (Figure~\ref{fig:tsne}F).
Consistent with the idea that NS1 is important for suppressing interferon, the one interferon-positive cell lacked detectable expression of the NS gene.
But other cells that lacked NS expression still failed to induce a detectable interferon response, despite often having $>$10\% of their mRNA derived from virus (Figure~\ref{fig:tsne}).
Therefore, most infected cells fail to express interferon even when the virus lacks its major interferon antagonist.
Our results are in line with other work showing that NS1-deficient influenza virus fails to deterministically induce interferon~\citep{Killip:2017ef,Kallfass:2013kp}.
Therefore, single cells infected at low MOI with a relatively ``pure'' stock of influenza virus only rarely activate innate-immune responses.

\subsection{Some host genes co-vary with viral gene expression.}
We examined whether any host genes were differentially expressed in cells with more viral mRNA.
We restricted this analysis to infected cells with all eight viral genes in order to focus on cellular genes that were associated with viral mRNA burden independent of effects due to the presence or absence of particular viral transcripts.
We identified 45 cellular genes that co-varied with viral gene expression at a false discovery rate of 0.1 (Figure~\ref{fig:cellulargenes}).

\begin{figure}
\centerline{\includegraphics[width=0.7\linewidth]{figures/p_cellular_heatmap.pdf}}
\caption{\label{fig:cellulargenes}
Cellular genes that co-vary in expression along with the the amount of influenza virus mRNA in individual cells expressing all eight viral genes.
Shown are all genes that are differentially expressed at a false discovery rate of 0.1. 
Genes for which the color goes from blue at left to red at right are expressed at higher levels in cells with more viral mRNA. 
\jdbcomment{scale shows?} }
\figdata{The full results of the differential expression test is in \texttt{p\_sig\_cellular\_genes.csv}.}
\figdata{The results of a gene-set analysis are in \texttt{p\_sig\_cellular\_genes.csv}.}
\figsupp[Many genes that co-vary with the amount of viral mRNA are involved in the response to oxidative stress. \label{figsupp:ROStable}]{
Table delineating genes in (Figure~\ref{fig:cellulargenes}) that are associated with the response to oxidative stress ~\citep{Duong:2017ju,Jung:2017jz,Lee:2017ba,Peuchant:2017hj,MacLeod:2016kt,Jiang:2016cf,Gorrini:2013jv,Miura:2013fr,Kim:2009gi,Banning:2005cz,Murray:2003ie,Doyle:1999ts}.
}{\includegraphics[width=0.7\linewidth]{figures/ROStable.pdf}}
\end{figure}

Many of the genes with increased expression in cells with more viral mRNA are known or suspected to be regulated by the Nrf2 master regulator in response to oxidative stress.
These genes produce proteins that are involved in detoxification of reactive oxygen species or resultant products, the management of misfolded proteins, the electron transport chain, or a general stress response (Figure~\ref{fig:cellulargenes}-Figure~supplement~\ref{figsupp:ROStable}). 
We additionally see reduced expression of the nitric oxide synthase interacting protein (NOSIP). 
Transient oxidative stress is known to occur during viral infection, and may act in a proviral fashion via MAPK activation driving vRNP export~\citep{Amatore:2014cs}.
The antioxidant response is thought to be largely antiviral, potentially through inhibition of MAPK activity \citep{Lin:2016ec,Sgarbanti:2014ht}.
Our data do not reveal whether the expression of genes involved in the response to oxidative stress are a cause or a symptom of higher levels of viral mRNA, and further investigation of this topic is an interesting area for future work.
Overall, the expression of some of the host genes in Figure~\ref{fig:cellulargenes}, along with inherent stochasticity and specific mutations to viral genes in individual cells, probably explain most of the variation in viral mRNA levels observed across the infected cells in our dataset.

\section{Discussion}
We have quantified the total transcriptome composition of single cells infected with influenza virus.
While we certainly observe a general increase in the amount of viral mRNA over time as expected from bulk measurements~\citep{Hatada:1989vz,Shapiro:1987ur}, there is wide variation in viral gene expression among individual infected cells.

One of the most striking forms of heterogeneity is the complete failure of some infected cells to express one or more viral genes, which we estimate occurs in about half the infected cells in our experiments.
The absence of viral genes in infected cells has been noted previously using more targeted approaches~\citep{Brooke:2013kb,Heldt:2015iz,Dou:2017cp}, and our work provides a holistic view by quantifying the the total viral transcriptional load as a function of the level of each mRNA.
We find that absence of any of the four genes that encode the viral RNP leads to a dramatic drop in the total amount of viral mRNA in an infected cell, consistent with prior bulk studies of viral transcription~\citep{Vreede:2004ip,eisfeld2015centre}.
Interestingly, the \emph{reason} some cells fail to express some viral genes remains unclear.
The prototypical influenza virion packages one copy of each of the eight gene segments~\citep{noda2006architecture,hutchinson2010genome}, but some virions surely package fewer~\citep{brooke2014influenza}.
However, it is also possible that much of the viral gene absence is due to stochastic loss of viral RNPs after infection but prior to the initiation of viral transcription in the nucleus.
Further work will be needed to distinguish between these possibilities.

 -- providing additional insight into the processes that drive influenza replication and its impact on the host at the cellular level.
The dataset we have generated is largely consistent with prior flow cytometry-based analyses, allowing us to build upon those prior observations with a more holistic view, beginning to understand the consequences of heterogeneity in influenza.
Overall we find that influenza replication is defined broadly by both significant stochasticity and significant robustness, and that a view of viral infection as either strictly deterministic or incredibly chaotic fails to accurately describe the infectious process.
Moreover we found that stress, likely induced by oxidative damage, appears to define the cellular response to infection.

It is immediately apparent that a model of influenza infection wherein there is equivalent accumulation of transcripts across successfully infected cells is far to facile to describe the true situation.
While infection certainly proceeds over time, with gradual increase in influenza transcript abundance, the level of transcripts expressed in cells varies wildly.
While some of these differences are likely due to temporal variation in binding and internalization, it will be interesting in follow-up experiments to determine whether there is an ultimate level of influenza transcript that is invariably reached or whether the variation we observe is indicative of a final setpoint.
Indeed, our analyses of host transcriptomes suggests that those cells expressing higher levels of influenza are distinguishable, even at 6 hours, from later time points expressing similar influenza burden, giving an indication that there are features above-and-beyond temporal variation driving differences in viral mRNA expression\jdbcomment{If we make this point we need to add the ranked graph to the results and address it} .

However, despite extensive heterogeneity, there seems to be robustness to the infectious lifecycle.
For example, influenza, regardless of the viral burden, exhibits relatively constrained ratios between the viral transcripts.
This suggests that the processes that are driving heterogeneity of influenza mRNA abundance act equivalently throughout the viral genome.
Potential mechanisms include components such as binding and entry kinetics that act on a virion and thus on all segments packaged therein, limitations of transcription rate that are dependent upon the same pool of cellular resources, or potentially even feedback-dependent mechanisms controlling viral transcription.

Perhaps even more important to influenza's success in the face of stochasticity is the capacity of the virus to control the host innate immune response despite a large number of non-infectious events. 
While non-infectious events might be considered metabolic 'waste', they would, theoretically, not impact the growth of successful viral infections. 
However, recognition of any virus, infectious or not, and production of interferon, would impact the entire viral population. 
While we cannot extrapolate much from our single interferon-positive cell, it can be stated that interferon induction by influenza is rare in our dataset. 
This is consistent with prior experiments using reporter cell lines,  which places the induction at ~1\% of cells with detectable influenza protein expression -- albeit this value is likely sensitive to differences in viral propagation and strain utilized ~\citep{Killip:2017ef, Chen:2010cr}. 
If the major influenza interferon antagonist, NS1, is absent in ~10-20\% of infections, and lowly expressed in those infections engaging in only primary transcription, we must posit that NS1 alone is insufficient to explain influenza evasion of innate immunity. 
Interestingly, if other segmented viruses also exhibit a pattern of stochastic segment loss, they too would need to overcome this hurdle and escape innate immunity through mechanisms relying on more than a gene product generated by a single segment.

In addition to viral mRNA kinetics, our analysis also allows us to observe those features of the cellular transcriptome that covary with influenza transcription.
Critically, our analysis does not require synchronization or infection with a high MOI, and thus avoids potential artifacts produced by either approach. 
We can therefore see extreme changes in a small subset of cells within a larger population.
Moreover, we can also observe how transcriptional changes partition between cells, and exclude dramatic changes such as interferon induction that, while important, can mask other interesting biologically-relevant responses. 
With this in mind, the predominant response associated with increased influenza mRNA transcript abundance appears to be the oxidative stress pathway. 
Generally characterized as antiviral, it is possible that cellular stress pathways represent an additional arm of innate immunity -- detecting a cellular state consistent with a loss of homeostasis as resources are diverted to viral propagation. 
Intriguingly, one element of the response, cyclophilin A (\emph{CypA}), is both thought to suppress influenza replication through degradation of matrix protein as well as act as a proinflammatory signal ~\citep{Liu:2009er, Jin:2004ii}.
It will be interesting to follow up on these observations and determine the broader mechanisms through which cellular stress pathways are involved in pro- or anti-viral responses.

The success garnered here underscores that single-cell approaches seem distinctly suited to study of viral pathogens, as not only are many host responses governed by bimodality and stochasticity ~\citep{Shalek:2014ey,Shalek:2013ej}, but also viral populations themselves exhibit a staggering array of diversity.
Our results with influenza have provided significant support for earlier analysis of viral stochasticity, further defined the sources and impacts of that stochasticity, and generated potentially fertile avenues of research. 
We stand now at an exciting moment in virology, as technology allows us to observe population-level measurements as emergent phenomena defined by underlying individual stochasticity rather than analyzing viral replication in a strictly deterministic framework.
It is our hope that this work will serve not only as a window into influenza replication, but as a part of a grander design in re-examining and understanding the fundamental features of viral infection.

MATERIAL FROM INTRODUCTION THAT MIGHT BE USEFUL HERE:

Recent advances in sequencing and visualization technologies have revealed that cell-to-cell heterogeneity is a fundamental feature of biological processes -- impacting events from bacterial entrance into quiescent states to organogenesis in multicellular organisms~\citep{JohnstonJr:2010jn}.
Due to this heterogeneity, biological pathways generated from bulk population data may not accurately describe the timing, order, or even the identity of events within single cells ~\citep{Trapnell:2014kg}.
Nearly no other feature of biological life is as defined by heterogeneity as viral infection, wherein events can be highly stochastic due to the incredible combination of both genetic and non-genetic sources of variation.

The contribution of heterogeneity to infectious processes has been well appreciated in the study of viral quasispecies, wherein pathogenesis is best described as an emergent process dependent upon underlying genetic diversity ~\citep{Lauring:2010if}.
This concept has revolutionized the study of RNA viruses: including polio, hepatitis C, and influenza \citep{Echeverria:2015ew, Kuroda:2010gt}. 
Direct evidence of the contribution of the quasispecies to pathogenesis has been observed -- for instance mutant strains of polio that produce less variation exhibit significantly less fitness in animal models of infection despite displaying wild-type replication kinetics \emph{in vitro}~\citep{Vignuzzi:2005ju}.
However, these studies have largely been limited to population-level analyses, which can link standing diversity to complex biological phenomena but are unable to deconvolute precise events in individual infected cells.

Non-genetic variation has also been described to contribute to the host response to infection. 
Single-cell transcriptomic analyses have revealed incredibly coordinated responses driven by stochasticity, even in tissue culture models using genetically homogeneous cells.
This heterogeneity in response has been observed both with respect to pure ligands such as lipopolysaccharide, wherein differences must be host-cell intrinsic ~\citep{Shalek:2014ey,Shalek:2013ej,Rand:2012kl}, and with respect to pathogens, wherein differences are likely a complex mixture of host-cell intrinsic and pathogen-specific heterogeneity ~\citep{Avraham:2015cj}. 
Moreover, stochasticity is observed both in interferon production, which drives the anti-pathogen state, and the expression of the downstream effector molecules of innate immunity.
As inflammation is a potentially deleterious process, restricting a response to a small subset of cells upon initial recognition of a threat might be particularly advantageous.

We now hope to apply new single-cell transcriptomic technologies to the study of influenza replication in a tissue culture model of infection.
Early single-cell transcriptomic approaches were laborious -- requiring the generation of sequencing libraries on individual cells~\citep{Tang:2009kj,ld:2012gj}.
Breakthroughs in microfluidics helped to automate this process, but retained strictly parallel processing of sequencing libraries.
More recent technologies have introduced more complex means of barcoding and partitioning cells, allowing for the generation of libraries consisting of thousands of cells with ?single pot? downstream processing.
These include the use of reverse-phase emulsions, gravity capture, and even \emph{in situ} reverse transcription after cellular partitioning~\citep{JunSSi,Gierahn:2017bc,Klein:2015ki,Macosko:2015ht}. 
We hope to leverage these technological advances in the study of influenza, using transcriptomic analysis to understand the phenotypic heterogeneity that abounds during infection with this virus.

\section{Methods and Materials}

\subsection{Cell lines and viruses}
The following cell lines were used in this study: the human lung epithelial carcinoma line A549 (ATCC CCL-185), the MDCK-SIAT1 variant of the Madin Darby canine kidney cell line overexpressing human SIAT1 (Sigma-Aldrich 05071502), and the human embryonic kidney cell line 293T (ATCC CRL-3216). 
All cells were maintained in D10 media (DMEM supplemented with 10\% heat-inactivated fetal bovine serum, 2 mM L-glutamine, 100 U of penicillin/ml, and 100 \si{\micro}g of streptomycin/ml) at 37 \si{\degreeCelsius} at 5 \% CO\textsubscript{2}.

Wildtype A/WSN/1933 (H1N1) influenza virus was generated by reverse genetics using the plasmids pHW181-PB2, pHW182-PB1, pHW183-PA, pHW184-HA, pHW185-NP, pHW186-NA, pHW187-M, and pHW188-NS~\citep{hoffmann2000dna}.
The sequences of the viral RNAs encoded in these plasmids are in Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}.
Reverse-genetics plasmids encoding the synonymously barcoded WSN virus were created by using site-directed mutagenesis to introduce two synonymous mutations near the 3' end of the mRNA for each viral gene.
The sequences of the synonymously barcoded viral RNAs are in Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}.

To generate viruses from these plasmids, we transfected an equimolar mix of all eight plasmids into cocultures of 293T and MDCK-SIAT1 cells seeded at a ratio of 8:1. 
At 24 hours post-transfection, we changed media from D10 to influenza growth media (Opti-MEM supplemented with 0.01\% heat-inactivated FBS, 0.3\% BSA, 100 U of penicillin/ml, 100  \si{\micro}g of streptomycin/ml, and 100 \si{\micro}g of calcium chloride/ml).
At 48 hours post-transfection we harvested the virus-containing supernatant, pelletted cellular material by centrifugation at 300 x g's for 4 minutes, and stored aliquots at of the clarified viral supernatant -80 \si{\degreeCelsius }. 
We then titered thawed aliquots of viral by TCID50 on MDCK-SIAT1 cells, computing titers via the formula of \citet{reed1938simple}.
To generate our ``high-purity'' stocks of viruses for the single-cell sequencing experiments, we then infected MDCK-SIAT1 cells at an MOI of 0.01, and let the virus replicate for 36 hours prior to harvesting aliquots that were again clarified by low-speed centrifugation, aliquoted, stored at -80  \si{\degreeCelsius }, and titered by TCID50.
The high-MOI passage (high-defective particle) stock used in Figure~\ref{fig:viruspopulations} was generated by instead passaging in MDCK-SIAT1 cells twice at an MOI of 1 for 48 hours.

For the experiments in Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, we created a virus that carried an HA gene segment in which GFP replaced most of the HA coding sequence, following a scheme first described by \citet{marsh2007specific}.
Briefly, we created a plasmid encoding a viral RNA with GFP in place of the HA coding sequence in the context of the pHH21~\citep{Neumann:1999ws} reverse-genetics plasmid, removing potential start codons upstream of the GFP (see Figure~\ref{fig:coexpression}-source~data~\ref{figdata:HAflank} for the sequence of the viral RNA).
We then generated GFP-carrying virus by reverse-genetics in cells constitutively expressing HA~\citep{Doud:2016gm}.
To obtain sufficient titers, this HA-eGFP virus was expanded for 44 rather than 36 hours after initiating infection at an MOI of 0.01.

\subsection{qPCR}
For the qPCR in Figure~\ref{fig:viruspopulations} and Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:cyclohexamide}, A549 cells were seeded at 3$\times$10\textsuperscript{5} cells / well in a 6-well tissue culture plate in D10 the day prior to infection. 
On the day of infection, a single well was trypsinized and the cells were counted in order to determine the appropriate amount of virus to use to achieve the intended MOI.
Immediately before infection, D10 was replaced with influenza growth media.
For cells incubated with cyclohexamide, the compound was added to a final concentration of 50  \si{\micro}g/ml at the time of infection -- previously confirmed to be sufficient to halt viral protein production~\citep{Killip:2014kz}.
RNA was purified using the QIAGEN RNeasy plus mini kit following manufacturer's instructions. 
cDNA was synthesized using an oligoDT primer and the SuperScript\textsuperscript{TM} III first-strand synthesis supermix from ThermoFisher using manufacturer's directions. 
Transcript abundance was measured using SYBR\textsuperscript{TM} green PCR master mix, using a combined anneal/extension step of 60 \si{\degreeCelsius } for one minute with the following primers: \emph{HA}: 5'-GGCCCAACCACACATTCAAC-3', 5'-GCTCATCACTGCTAGACGGG-3', \emph{IFNB1}: 5'-AAACTCATGAGCAGTCTGCA-3', 5'-AGGAGATCTTCAGTTTCGGAGG-3', \emph{L32}: 5'-AGCTCCCAAAAATAGACGCAC-3', 5'-TTCATAGCAGTAGGCACAAAGG-3'.   
Biological triplicates were performed for all samples.

For the measurements of viral genomic HA content in Figure~\ref{fig:viruspopulations}A, vRNA was harvested from 80 \si{\micro}l of viral supernatant by the addition of 600 \si{\micro}l of RLT plus before proceeding with the standard QIAGEN RNeasy Plus Mini kit protocol.
The cDNA was generated using SuperScript\textsuperscript{TM} III first-strand synthesis supermix using the manufacturer's protocol, and using the universal vRNA primers of \citet{Hoffmann:2001vj} with the modifications described in  \citet{Xue:2017dl}.
The qPCR was then performed as for mRNA measurements.
A standard curve was generated from three independent dilutions of the HA-encoding reverse genetics plasmid. 
All vRNA values represent three independent RNA extractions with two replicate qPCR measurements.

\subsection{Flow cytometry titering and analyses}
To determine viral titers in terms of HA-expressing units and for the flow cytometry in Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, A549 cells were seeded in a 6-well plate and infected as described above for the qPCR analyses.
Cells were harvested by trypsinization, resuspended in phosphate-buffered saline supplemented with 2\% heat-inactivated FBS, and stained with 10 \si{\micro}g/ml of H17-L19, a mouse monoclonal antibody confirmed to bind to WSN HA in a prior study~\citep{Doud:2017bw}.
After washing in PBS plus 2\% FBS, the cells were stained with a goat anti-mouse IgG antibody conjugated to APC.
Cells were then washed, fixed in 1\% formaldehyde, and washed further before a final resuspension and analysis. 
We then determined the fraction of cells that were HA positive and calculated the HA-expressing units assuming a Poisson infection process.

For Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, after gating to exclude multiplets in FlowJo, data were extracted using the R package flowCore~\citep{LeMeur:2007uo} and analyzed using a custom Python script.
Guassian kernel density estimates were obtained using the scipy stats package method, guassian\_kde, using automatic bandwidth determination~\citep{vanderWalt:2017dp}.

\subsection{Infections for single-cell mRNA sequencing}
Single-cell sequencing libraries were generated using the 10x Chromium Single Cell 3' platform~\citep{zheng2017massively} using the V1 reagents.

All time points except for the 8-hour replicate were prepared on the same day.
For the infections, A549 cells were seeded in a 6-well plate, with two wells per time point. 
A single well of cells was trypsinized and counted prior to initiation of the experiment for the purposes of calculating MOI.
Wild-type and synonymously barcoded virus were mixed to an estimated ratio of 1:1 based on prior, exploratory, single-cell analyses (data not shown). 
At the initiation of our experiment, the wells for all time points were changed from D10 to influenza growth media.
Cells were then infected with 0.3 HA-expressing units of virus per cell (a determined by flow cytometry).
The infections were performed in order of time point: first the 10-hour time point, then the 8-hour, and then the 6-hour time point.
At one hour after infection, the media for each time point was changed to fresh influenza growth media.
Note that the HA-expressing units were calculated without this additional washing step, and so likely represent an overestimate of our final infectious dose (consistent with the fact that fewer than 30\% of cells appear infected in the single-cell sequencing data).
All cells were then harvested for single-cell analysis concurrently -- ensuring all had spent equivalent time in changed media .
For our replicate 8-hour time point, cells were infected as above except that the cells were infected at 0.1 HA-expressing units of virus per cell but no wash step was performed, and the sample was prepared on a different day.
After harvest, cells were counted using disposable hemocytometers and diluted to equivalent concentrations with an intended capture of 3000 cells/sample following the manufacturer's provided by 10x Genomics for the Chromium Single Cell platform.
All subsequent steps through library preparation followed the manufacturer's protocol.
Samples were sequenced on an Illumina HiSeq. 
The deep sequencing data are available on the Sequence Read Archive under accession \jdbcomment{needs to be added.}

\subsection{Computational analysis of single-cell mRNA sequencing data}
Jupyter notebooks that perform all of the computational analyses are available at \url{https://github.com/jbloomlab/flu_single_cell} \jdbcomment{This GitHub repository will be made public upon acceptance of the manuscript. If you are a reviewer who needs access, please contact the editor.}, and are also available in Supplementary~file~\ref{suppfile:code}.

Briefly, the raw deep sequencing data were processed using the 10X Genomics software package CellRanger (version 2.0.0).
The reads were aligned to a concatenation of the human and influenza virus transcriptomes.
The human transcriptome was generated by filtering genome assembly GRCh38 for protein coding genes defined in the GTF file GRCh38.87.
The influenza virus transcriptome was generated from the reverse-complement of the wildtype WSN viral RNA sequences as encoded in the reverse-genetics plasmids (Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}).
CellRanger calls cells based on the number of observed cell barcodes, and creates a cell-gene matrix.
We used custom Python code to annotate the cells in this matrix by the number of viral reads that could be assigned to the wildtype and synonymously barcoded virus.
This annotated cell-gene matrix is in Supplementary~file~\ref{suppfile:cellgenematrix}.
A Jupyter notebook that performs these analyses is in Supplementary~file~\ref{suppfile:code}.

The annotated cell-gene matrix was analyzed in R, primarily using the Monocle package (version 2.4.0)~\citep{Qiu:2017dw,Trapnell:2014kg}.
A Jupyter notebook that performs these analyses is in Supplementary~file~\ref{suppfile:code}.
For each sample, cell barcodes that had $>$2.5-fold fewer or more UMI counts mapping to cellular transcripts than the sample mean were excluded from downstream analyses (see red vertical lines in Figure~\ref{fig:cells}C).
In order to determine an appropriate cutoff for how many reads a cell needed to contain in order to be classified as infected, we calculated the mean viral barcode purity across all cells that contained at least a given fraction of viral mRNA and had multiple viral reads that could be assigned a barcode (Figure~\ref{fig:viralbarcodes}B,C).
We then determined the threshold fraction of viral mRNA at which the mean purity no longer increased as a function of the amount of viral mRNA.
This threshold represents the point at which we have effectively eliminated cells that have low barcode purity simply due to lysis-acquired reads sampled randomly from both viral barcodes.
The 10-hour sample has the most viral mRNA, and so is expected to have the most lysis-associated influenza reads.
We therefore determined the threshold using just this sample (Figure~\ref{fig:viralbarcodes}C), and then applied this threshold to all samples.
This procedure is expected to be conservative, and may miss some truly infected cells with very low amounts of viral mRNA.
For subsequent analyses, we retained all infected cells and a subsample of uninfected cells (the greater of 50 or the number of infected cells for that sample).
The rationale for subsampling the uninfected cell is that the vast majority of cells are uninfected, and we did not want these cells to completely dominate the downstream analyses.

For the semi-supervised t-SNE clustering, we used Monocle's cell hierarchy function to bin cells into those with no viral mRNA, <2\% viral mRNA, between 2\% and 20\% viral mRNA, and >20\%. 
Candidate marker genes for t-SNE dimensionality reduction were then determined using the Monocle function markerDiffTable, excluding the effects of sample variation and the number of genes identified in a given cell, using a q-value cutoff of 0.01.
The specificity of these markers was determined using the function calculateMarkerSpecificity -- the top 40 markers were retained, and used to place populations in a two-dimensional plane based on tSNE dimensionality reduction.

For the analyses of cellular genes that differed in expression as a function of the amount of viral mRNA, we only considered cells that expressed all 8 viral mRNAs to avoid effects driven simply by viral gene absence.
We also only considered cellular genes in the differential gene analysis, since viral gene expression will tautologically co-vary with the amount of viral mRNA.
Additionally, because influenza virus has the capacity to degrade or prevent the synthesis of host mRNAs~\citep{BercovichKinori:2016iw} and contributes significantly to the total number UMIs in some cells, we calculate size factors (a scalar value representing efficiency of UMI capture) based on cellular transcripts alone. 
Finally, we assigned all cells a ceiling fraction of mRNA from virus of 25\% so that a few extremely high-expressing cells did not dominate. 
Cellular genes with expression that co-varied with the fraction of viral mRNAs in a cell were then determined using the Monocle differentialGeneTest, after removing variance explained by sample to sample variation. 
Figure~\ref{fig:cellulargenes} shows all genes that were significantly associated with the fraction of mRNA from virus at a false discovery rate of 0.1.

\section{Acknowledgments}
We thank the Fred Hutch Genomics Core for performing the Illumina deep sequencing.

\nocite{*} % This command displays all refs in the bib file
\bibliography{references}

\clearpage

\begin{suppfile}
\caption{\label{suppfile:code}
Computer code for the analyses.
This ZIP file contains a Jupyter notebook that runs CellRanger to align and annotate the reads, and a Jupyter notebook that uses Monocle to analyze the cell-gene matrix.
The ZIP file also includes associated custom scripts.}
\end{suppfile}

\begin{suppfile}
\caption{\label{suppfile:cellgenematrix}
The annotated cell-gene matrix in Matrix Market Format.}
\end{suppfile}

\end{document}

\end{document}